% Los objetivos específicos de cada uno de los subproyectos participantes, enumerándolos brevemente, con claridad, precisión y de manera realista (acorde con la duración prevista del proyecto).%% En los subproyectos con dos investigadores principales, deberá indicarse expresamente de qué objetivos específicos se hará responsable cada uno de ellos.%

\subsubsection*{Objectives of the ENG subproject}

The ENG subproject centralises the front-end electronics, DAQ, and slow controls of the NEW and NEXT-100 detectors. It is coordinated by the UPV.

The specific objectives of this sub-project (also called NEXT projects or NP) are:

\begin{enumerate}
\item {\bf MI (Mechanical Infrastructures)}: This includes the construction and commissioning of the working platform, seismic pedestal and lead castle. This NP is coordinated by Prof. Jose Luis P\'erez (UPV). 
\item {\bf FEE (Front End Electronics)}: Design, fabrication and commissioning of the front-end electronics for the PMTs and the SiPMs for NEW and NEXT-100. The NP leader is the co-PI of the subproject, Prof. Francisco Toledo (UPV).
\item {\bf DAQ}: Design, fabrication and commissioning of the data acquisition modules for NEW and NEXT-100. The NP leader is the second co-PI of the subproject, Prof. Raul Esteve (UPV).
\item {\bf Slow control}: Design, fabrication and commissioning of the slow control for NEW and NEXT-100. The project leader is technical engineer Vicente Álvarez.
\item {\bf Online}: Design and commissioning of the online monitoring for NEW. Interfaces with offline, DAQ and Slow Control. The project leader is 
informatics engineer Toni Mar\'i.
\end{enumerate}

