\subsubsection*{Objectives of this coordinated project and the NEXT Project Management Plan}

The objectives of this coordinated project are described in the NEXT Project Management Plan (PMP) and correspond to specific Working Packages (WP) defined there. The WP are:

\begin{itemize}
\item {\bf WP1: PV (Pressure Vessel)}: Construction of the NEW and NEXT-100 pressure vessels. 
\item {\bf WP2: GS (Gas System)}: Installation and commissioning of the Gas System for NEW and NEXT-100 at the LSC. 
\item {\bf WP3:IS (Infrastructures)}: Construction of the infrastructures (working platform, seismic pedestal, lead castle, radon suppression system and clean tent), needed for the operation of NEW/NEXT-100.
\item {\bf WP4: FC (Field Cage)}: Construction of the field cage, light tube, HVFT and EL grids of NEW and NEXT-100.
\item {\bf WP5: EP (Energy plane)}: Construction of the NEW/NEXT-100 energy planes.
\item {\bf WP6: TP (Tracking plane)}: Construction of the NEW/NEXT-100 tracking planes.
\item {\bf WP7: FEE (Front-End electronics)}: Front-End electronics of energy and tracking planes.
\item {\bf WP8: DAQ (Data Acquisition )}: Data  acquisition of the energy and tracking planes.
\item {\bf WP9 ONL (Online) }: Online monitoring for NEW/NEXT-100. 
\item {\bf WP10: SLW (Slow Controls)}: Slow controls of the detectors, gas system and ancillary systems.
\item {\bf WP11: MC (Monte Carlo)}: Monte Carlo simulation of the detector(s) and implementation of the background model. 
\item {\bf WP12: REC (Reconstruction)}: Offline reconstruction of events in NEW/NEXT-100.
\item {\bf WP13: SOF (Software)}: Development and maintenance of software tools (reconstruction and analysis frameworks, release tools, software libraries) and coordination of data and Monte Carlo productions. 
\item {\bf WP14: CAL (Calibration)}: Calibration of the NEW/NEXT-100 with radioactive sources. 
\item {\bf WP15: RAD (Radiopurity)}: Screening, using the LSC facilities, as well as several other techniques, of all the components being used in the  NEW/NEXT-100 detectors. 
\end{itemize}

The COORD subproject has two set of objectives: the first set (construction, commissioning and integration, or CCI) objectives are co-ordinated by the PI-1, and the second set (software and analysis, SOA) are co-coordinated by the PI-2 of the subproject. The CCI objectives are those included in WP1 to WP6. The SOA objectives are those included in WP11, and WP13. The objectives of the ENG subproject are those included in WP7, WP8, WP9 and WP10. The objectives of the CALREC subproject are those included in WP12 and WP14. Finally, the objectives of the RPUR subproject are those included in WP15.  


